---
title: "Code implementation of bayesian method"
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---


```{r, include=FALSE}
source("code/function.R")
```
<h1> Introduction </h1>
To implement the proposed method from the home page we've created the bayesian_method function. We have the following input arguments and output:
<br>
**Usage:**

bayesian_method(cover_data, presence_absence_data, n = 16 , remove_column = 0)

**Arguments:**

* **Cover_data**: Data frame with plots as rows and species as columns. Each cell in the data frame contains cover data for a given specie in a given plot.
* **presence_absence_data**: Data frame with plots as rows and species as columns. Each cell in the data frame contains a 1 if the specie is present in the plot or a 0 if the specie is absent in the plot.
* **n**: An integer. A number each cell in the cover data gets divided by if the cover data is not in the interval $[0,1]$. In the NOVANA dataset n = 16 secures this property. If the cover data is already on the interval $[0,1]$, one can just let n = 1. n is the number of pins used in each blot, so by dividing each entry with n, we normalize the data.
* **remove_column**: An integer. In case the first columns in the cover_data and presence_absence_data contain data not about the species but rather information on the plots, these columns have to be removed. The numbers of columns to be removed must be the same in the cover_data and presence/absence_data. The removed columns will be included again in the  output. In the NOVANA dataset the first columns are "site", "plot", year" so remove_column = 3 when this datasat is applied.

**Output** 

* The function returns a dataframe with plots as rows and species as columns, and where each entry contain new abundance data for a specie in a plot.

<h1>Description of code </h1>
First, two subsets of the input datasets are created without the columns that need to be removed. These are called cover_data and freq_data. Next we loop through the species in the cover_data and remove the species that has no presence in any plot in the freq_data. Both the input cover dataset and the cover_data are updated accordingly. This way we get rid of some of the data that is of no use, which will make the function run faster. 
<br><br>
A dataframe "beta_fit" is created. This dataframe will contain a column for each species in the cover data and corresponding $a$ and $b$ values from the fitted beta-distribution for the prior distribution of the species. Every cell in the cover_data is divided with the input constant $n$. This ensures that every cell is on the interval $[0,1]$ as this is the interval on which the beta distribution is defined. In the NOVANA dataset we let $n = 16$, because the total number of pins used in each plot is 16.
<br>
```{r, eval = FALSE}
 library(fitdistrplus)
  cover_data <- cover[,(remove_column+1):ncol(cover)]
  freq_data <- presence_absence_data[,(remove_column+1):ncol(presence_absence_data)]


  for (species in colnames(cover_data)){
    if (sum(freq_data[,species]) == 0){
      cover = cover[,!(names(cover) %in% species)]
      cover_data <- cover_data[,!(names(cover_data) %in% species)]
    }
  }
```

As discussed on the home page we will fit a beta-distribution as the prior distribution for each species using all the cover data where the correspoding presence/absence data has a 1. This is saved in the beta_data vector.
<br>
We only fit a beta-distribution to species that have more than 1 unique cover data value. To do the fitting we use the package "fitdistrplus" with method = "method of moments" because this method can handle the values of 0 and 1. In some few special cases we can have a species that is only present in one plot. For these species we let the $a$ and $b$ values be zero because we cannot fit a prior beta distribution for these. Later in the code when we calculate the updated beta distribution, the prior will have no effect on this posterior distribution, hence the posterior will just be made by the coverdata, since this is the likelohood.
<br>
```{r, eval = FALSE}
#For each species we save the shape parameters from the fitted prior beta distribution in a data frame.
  beta_fit <- data.frame(matrix(ncol = 3, nrow = 0))

  colnames(beta_fit) <- c("species","a", "b")

  for (name in colnames(cover_data)) {
    species <- cover_data[,name]/n

    #Remove all plots with 0 in presence/absence data
    beta_data <- species[freq_data[[name]] == 1]


    if (length(unique(beta_data)) > 1) {
      beta_data_fitted <- fitdist(beta_data, "beta", method = "mme")
      beta_fit[nrow(beta_fit) + 1,] <- c(name, beta_data_fitted$estimate[1], beta_data_fitted$estimate[2])

    }
    else {
      beta_fit[nrow(beta_fit) + 1,] <- c(name, 0,0)

    }
  }

```

After creating this dataframe of prior distributions, we can now combine the cover data and presence/absence data to get a new abundance estimate for each species in each plot. 
<br><br>
We have an outer loop where we loop through each plot in the cover_data. In each plot we want the species that are spotted in this plot. However, some species have a 1 in the presence/absence data but no data in the cover data. We have decided to exclude these species. This is done with use of the setdiff() function. After this is done, we loop trough the spotted species in the plot and create the new abundance estimate. These are saved in the corresponding cells in the input cover dataset. Because of some numerical issues, we let the abundance estimate be zero if the mean_posterior is below some threshold that we have set at 0.00001.
<br>
```{r, eval = FALSE}
  #For each cell in the input cover data we find the new abundance estimate and save it to this data frame
  for (plot in 1:nrow(cover_data)) {

    #For each row we find the species that has a 1 in the presence/absence data
    species_spotted_in_freq <- colnames(freq_data[c(freq_data[plot,]  == 1)])

    not_in_cover <- setdiff(species_spotted_in_freq,colnames(cover_data))

    species_spotted_in_freq <- setdiff(species_spotted_in_freq, not_in_cover)


    #We calculate new abundance estimate for each spotted species in the plot
    for (species_spotted in species_spotted_in_freq ) {


      alpha_post <- as.numeric((as.numeric(beta_fit[beta_fit$species == species_spotted,]$a) +
                                  as.numeric(cover_data[[species_spotted]][plot]) ))
      beta_post <-  as.numeric(beta_fit[beta_fit$species == species_spotted,]$b) + n - as.numeric(cover_data[[species_spotted]][plot])

      mean_posterior <- (alpha_post)/(alpha_post+beta_post)

      if(mean_posterior < 0.00001){
        cover[plot,species_spotted] <- 0
      }
      else{
        cover[plot,species_spotted] <- mean_posterior * n
      }

    }
  }
```
<br>
Before we return the dataframe with the new abundance estimates, we remove all the columns (species) that have a zero in all plots. This makes output dataset lesser and thereby easier to work with in the following analysis. If one wants to include the these species, the for-loop can just be omitted.
```{r, eval = FALSE}
  for (species in colnames(cover_data)){
    if (sum(cover[,species]) == 0){
      cover = cover[,!(names(cover) %in% species)]
    }
  }

  return(cover)
```

<h1>Full code</h1>
Below is the entire function in one code chunk
```{r}
bayesian_method
```

