---
title: "Code implementation of Bayesian model"
output:
  workflowr::wflow_html:
    toc: true
    css: style.css
editor_options:
  chunk_output_type: console
---


```{r, include=FALSE}
source("code/function.R")
```

The implementation of the function "beta_binomial_cover_update" has the following properties. 

**Usage:**

`beta_binomial_cover_update(cover_data, presence_absence_data, n = 16 , remove_column = 0)`

**Arguments:**

* **cover_data**: A data frame with plots as rows and species as columns. Each cell in the data frame contains cover data for a given species in the small square of a given plot.
* **presence_absence_data**: A data frame with plots as rows and species as columns. Each cell in the data frame contains a 1 if the species is present in the plot or a 0 if the species is absent in the plot.
* **n**: An integer. A number each cell in the cover data gets divided by if the cover data is not on the interval $[0,1]$. In the NOVANA dataset n = 16 secures this property. If the cover data is already on the interval $[0,1]$, one can just let n = 1.  
* **remove_column**: An integer. In case the first columns in the cover_data and presence_absence_data contain data not about the species but rather information on the plots, these columns have to be removed. The numbers of columns to be removed must be the same in the cover_data and presence/absence_data. The removed columns will be included again in the  output. In the NOVANA dataset the first columns are "site", "plot", year" so remove_column = 3 when this dataset is applied.

**Output** 

* The function returns a data frame similar to the cover data but now containing new cover data estimated using the method outlined on the [Model](model.html) page.


<h1>Description of code </h1>
First, two subsets of the input datasets are created without the columns that need to be removed. These are called cover_data and freq_data. We loop through the species in the cover_data and remove the species that has no presence in any plot in the freq_data. Both the input cover dataset and the cover_data are updated accordingly. This way we get rid of some of the data that is of no use, which will make the function run faster. 

```{r, eval = FALSE}
 library(fitdistrplus)
  # We remove the columns that are not wanted, this is done by only selecting
  # the columns from after the column number that the user want to remove. This is done for both
  # the cover data and the present/absence data
  cover_data <- cover[,(remove_column+1):ncol(cover)]
  freq_data <- presence_absence_data[,(remove_column+1):ncol(presence_absence_data)]

  # We now remove the columns in cover data that contain a species not present in any of the plots.
  # This is done by finding the species not present in any plot in the presence/absence data, and then
  # removing the corresponding columns from the cover data.
  for (species in colnames(cover_data)){
    if (sum(freq_data[,species]) == 0){
      cover = cover[,!(names(cover) %in% species)]
      cover_data <- cover_data[,!(names(cover_data) %in% species)]
    }
  }
```

A data frame "beta_fit" is created. This data frame will contain a column for each species in the cover data and corresponding $a$ and $b$ values from the fitted beta-distribution for the prior distribution of the species. Every cell in the cover_data is divided with the input constant $n$. This ensures that every cell is on the interval $[0,1]$ as this is the interval on which the beta distribution is defined.


We only fit a beta-distribution to species that have more than 1 unique cover data value. To do the fitting we use the package "fitdistrplus" with method = "method of moments" because this method can handle the values of 0 and 1. In some few special cases we can have a species that is only present in one plot. For these species we let the $a$ and $b$ values be zero because we cannot fit a prior beta distribution for these. Later in the code when we calculate the updated beta distribution, the prior will have no effect on this posterior distribution
<br>
```{r, eval = FALSE}
  # We then create a data frame to hold the fitted parameter values for the prior distribution
  # for each species
  beta_fit <- data.frame(matrix(ncol = 3, nrow = 0))
  
# We name the columns in the new data frame
  colnames(beta_fit) <- c("species","a", "b")

  # For each species we calculate the parameters for the fitted prior beta distribution and save them
  # in the data frame, beta_fit.
  for (name in colnames(cover_data)) {
    # We divide by n, to normalize the data, so each entry in the cover data, contains values between
    # 0 and 1.
    species <- cover_data[,name]/n

    # We remove all plots with 0 in frekvens for the given specie.
    beta_data <- species[freq_data[[name]] == 1]


    # We now want to fit a prior beta distribution for the given species We can only fit a beta
    # distribution if there is more than 1 unique value, and therefore we use the argument
    # length(unique(beta_data)) > 1. If there is not more than 1 unique value, we make the parameters
    # of the distribution 0. Else we use method og moment "mme" to fit a prior beta distribution.
    if (length(unique(beta_data)) > 1) {
      # Here we fit a beta distribution
      beta_data_fitted <- fitdist(beta_data, "beta", method = "mme")
      # We then save the parameters in the data frame
      beta_fit[nrow(beta_fit) + 1,] <- c(name, beta_data_fitted$estimate[1], beta_data_fitted$estimate[2])

    }
    else {
      beta_fit[nrow(beta_fit) + 1,] <- c(name, 0,0)

    }
  }

```

After creating this data frame of prior distributions, we can now combine the cover data and presence/absence data to get a new abundance estimate for each species in each plot. 
<br><br>
We have an outer loop where we loop through each plot in the cover_data. In each plot we want the species that are spotted in this plot. However, some species have a 1 in the presence/absence data but no data in the cover data. We have decided to exclude these species. This is done with use of the setdiff() function. After this is done, we loop trough the spotted species in the plot and create the new cover estimate. These are saved in the corresponding cells in the input cover dataset. Because of some numerical issues, we let the abundance estimate be zero if the mean_posterior is below some threshold that we have set at 0.00001.
<br>
```{r, eval = FALSE}
  #For each cell in the input cover data we find the new abundance estimate and save it to this data frame.
  for (plot in 1:nrow(cover_data)) {

    #For each row (plot) we find the species that has a 1 in the presence/absence data.
    species_spotted_in_freq <- colnames(freq_data[c(freq_data[plot,]  == 1)])

    # We then find the species not in cover but is in the presence/absence data.
    not_in_cover <- setdiff(species_spotted_in_freq,colnames(cover_data))

    # We then choose the species that are both in cover and presence absence data.
    species_spotted_in_freq <- setdiff(species_spotted_in_freq, not_in_cover)


    #We calculate new abundance estimates for each spotted species in the plot
    for (species_spotted in species_spotted_in_freq ) {

      # We calculate the parameters of the posterior beta distribution for each species in the plot
      alpha_post <- as.numeric((as.numeric(beta_fit[beta_fit$species == species_spotted,]$a) +
                                  as.numeric(cover_data[[species_spotted]][plot]) ))
      beta_post <-  as.numeric(beta_fit[beta_fit$species == species_spotted,]$b) + n - as.numeric(cover_data[[species_spotted]][plot])

      # We find the mean value of the posteriro beta distribution for each specie in the plot,
      # with the parameters calculated above.
      mean_posterior <- (alpha_post)/(alpha_post+beta_post)

      # If the value is to small, we will change it to 0, else we time the value with n,
      # to unnormalise the value
      if(mean_posterior < 0.00001){
        cover[plot,species_spotted] <- 0
      }
      else{
        cover[plot,species_spotted] <- mean_posterior * n
      }

    }
  }
```
<br>
Before we return the data frame with the new cover estimates, we remove all the columns (species) that have a zero in all plots. This makes the output smaller and thereby easier to work with in the following analysis. If one wants to include the these species, the for-loop can just be omitted.
```{r, eval = FALSE}
#We then remove the columns from the cover data set, if the species is not in any of the plots.
  for (species in colnames(cover_data)){
    if (sum(cover[,species]) == 0){
      cover = cover[,!(names(cover) %in% species)]
    }
  }

# We return the updated cover data set which will be referenced to as beta binomial cover
  return(cover)
```

<h1>Full code</h1>
Below is the entire function in one code chunk
```{r}
beta_binomial_cover_update
```

