---
title: "Documentation for different ways of calculating diversity"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: TRUE
---

```{r, include=FALSE}
source("code/function_shannon.R")
```

Functions for calculating a simpler shannon index and species richness



<h1>Row_shannon </h1>



The simpler shannon index only uses cover data.

**Usage:**

row_shannon(cover_data ,species_indicator = NULL)

**Arguments:**

* Cover_data: A data frame with plots as rows and the cover data for each specie as columns.

* species_indicator: A string. Is used to specify which columns contain species data. It will use all columns which contain the string as a substring.  If left empty all columns in cover_data will be assumed to be for species.

```{r, echo = FALSE}
row_shannon
```


 
<h1>species_richness </h1>

Species richness is the number of observed species in a given plot.



**Usage:**

species_richness(frequency_data ,species_indicator = NULL)

**Arguments:**

* frequency_data: A data frame with plots as rows and the frequency data for each specie as columns (1 if the species is spotted, 0 if not).

* species_indicator: A string. Is used to specify which columns contain species data. It will use all columns which contain the string as a substring.  If left empty all columns in frequency_data will be assumed to be for species.

```{r, echo = FALSE}
species_richness
```
