---
title: "A bayesian method for better plant diverity estimates"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---
```{r, include=FALSE}
library(tidyverse)
library(kableExtra)
```

<h1>Introduction</h1>
Many methods for estimating the abundance of species within an area has been proposed in the litterature. In this vignette we will propose a new method with this goal where we combine cover data and prensence/absence data. We will be working with a subset of the NOVANA dataset as an example to show how this new method can be applied.
<br><br>
The NOVANA dataset consists of several parts. We will be using data of cover and presence/absence of species within each plot and also further data about each plot. 
<br>
Cover data of species in a given area can be done in many ways. In the NOVANA dataset the pin-point method is used to estimate the abundance of species in a plot. In the center of each plot $n$ pins is placed and the number of pins each specie is hit by is recorded. In the NOVANA dataset $n$ = 16 pins are used. As an example the below dataframe shows what cover data could look like
```{r, include=FALSE}
Cover <- data.frame(   a1 = c(2,0,5,0,1,0,16),
                       a2 = c(0,0,0,2,0,1,10),
                       a3 = c(0,0,1,1,16,0,2),
                       a4 = c(1,0,10,9,0,0,7),
                       a5 = c(0,0,0,3,0,0,0),
                       a6 = c(3,0,11,0,14,1,13),
                       a7 = c(0,0,7,4,0,1,0),
                       a8 = c(0,0,16,9,0,0,9),
                       a9 = c(2,0,0,1,11,0,5),
                       a10 = c(0,0,0,1,13,0,0))


rownames(Cover) <- c("Plot 1", "Plot 2", "Plot 3", "Plot 4", "Plot 5", "Plot 6", "Plot 7")
colnames(Cover) <- c("Specie 1", "Specie 2", "Specie 3", "Specie 4", "Specie 5", "Specie 6", "Specie 7", "Specie 8", "Specie 9", "Specie 10")
```

```{r}
kable(Cover, "html") %>% kable_styling()
```

The other part we will use to estimate the abundance of species in each plot is the precense/absence data. Within each plot it is recorded whether a specie is to be found somewhere in the plot. This is indicated by a binary variable with found = 1. Below is an example of what this data could look like
```{r, include=FALSE}
PersenceAbsence <- data.frame(   a1 = c(1,0,1,0,1,0,1),
                                 a2 = c(1,1,0,1,0,1,1),
                                 a3 = c(0,0,1,1,1,0,1),
                                 a4 = c(1,0,1,1,0,0,1),
                                 a5 = c(0,0,0,1,0,0,1),
                                 a6 = c(1,0,1,0,1,1,1),
                                 a7 = c(1,0,1,1,0,1,0),
                                 a8 = c(0,0,1,1,0,0,1),
                                 a9 = c(1,0,0,1,1,0,1),
                                 a10 = c(0,0,0,1,1,0,1))


rownames(PersenceAbsence) <- c("Plot 1", "Plot 2", "Plot 3", "Plot 4", "Plot 5", "Plot 6", "Plot 7")
colnames(PersenceAbsence) <- c("Specie 1", "Specie 2", "Specie 3", "Specie 4", "Specie 5", "Specie 6", "Specie 7", "Specie 8", "Specie 9", "Specie 10")
```


```{r}
kable(PersenceAbsence, "html") %>% kable_styling()
```

Here we also note that a specie can have a 1 in the presence/absence data but a 0 in the cover data as it is also the case in the above examples. We will make use of this property of the data in our method.

<h1>Theoretical discussion </h1>

The plant diversity is often calculated by only using the presence/absence data. However, in this method we loose a lot of valuable information from the cover data. In the following we propose a method that obtain an estimate of a specie's abundance in each plot that takes into account the spatial aggregation of the species which we can estimate using the cover data. This should also lead to better diversity estimates. We will do this by applying a bayesian framework and use a beta-binomial model. From the general setup in bayesian statistics we get that
$$
P(\theta|Y=y) \propto P(Y=y|\theta)\cdot \pi(\theta)
$$
where $P(Y=y|\theta)$ is the likelihood function and $\pi(\theta)$ is the prior distribution. We are interested in the posterior distribution $P(\theta|Y=y)$ which we can use to estimate the abundance of a specie in a plot. 
<br><br>
In the cover data each specie is hit by $y$ pins out of $n$ possible in each plot. A natural distribution to model this is the binomial distribution
$$
P(Y = y| \theta) = {n \choose y}\theta^y(1-\theta)^{n-y}
$$
<br>
We use the binomial model as the likelihood function in the bayesian framework that we apply.
<br>
We can create a prior distribution $\pi(\theta)$ for each specie by using both cover and presence/absence data. This is really the key idea in this method because this makes it possible to get a prior belief of a species spatial aggregation in a plot when we know that the specie is present.
<br><br>
The beta model is a conjugacy prior for the binomial model and, thus, is ideal to use. The beta-distribution is given by 
$$
\pi(\theta) = \text{beta}(\alpha, \beta) = \frac{\Gamma (\alpha+\beta)}{\Gamma (\alpha) \Gamma(\beta)}\theta^{\alpha -1}(1-\theta)^{\beta-1}
$$
When working with the bayesian framework, one has to decide on the paramters $\alpha$ and $\beta$. However, this can be difficult and instead we apply the empirical bayes method: For each specie we fit a beta-prior using all the cover data where the corresponding presence/absence data is 1. 
<br><br>
From the likelihood function and the prior distribution we get that the posterior will be an updated beta distribution for each specie in each plot
$$
P(\theta | Y = y) \sim Beta(\alpha + y, \beta + n-y)
$$
When the posterior distribution is found for a specie in a given plot, we use the mean posterior as a measure of the abundance for the specie in the plot. <br>

The mean posterior for a beta-distribution is given by:

$$
E\Big(P(\theta|Y =y)\Big) = \frac{\alpha + y}{\alpha + y+\beta + n-y}
$$

From the above theoretical discussion we get a new dataset with updated abundances for each specie in each plot. The next step for users of this method will be, as an xample, to apply some diversity measures for the plots in the new updated dataset. Some of commonly used diversity measures are described in [link](Shannon.html) and will be used in the examples below.

<h1> Implementation </h1>
```{r, include=FALSE}
source("code/function.R")
```

The above theoretical discussions of the beta-binomial model are implemented in the function "bayesian_method". 

**Usage:**

bayesian_method(cover_data, present/absence_data, n = 16 , remove_column = 0)

**Arguments:**

* **Cover_data**: Cover data in a data frame, with plots as rows and species as columns.
* **present/absence_data**:  present/absent data in a data frame, with plots as rows and species as columns. 1 for present and 0 for absent.
* **n**: If cover data is not between 0 and 1, what should it be divide with to be in the interval $[0,1]$.
* **remove_column**: Exclude first n column from the data analysis. In the NOVANA dataset this we would remove the first 3 columns as they include data about the plot, site and year.

<p></p>
The function returns a dataframe with updated abundance data.
<br><br>
For further specifications of the function click [link](function_shannon.html)


<h1>An example </h1>

We work with data only including the terhabtype = 6230 from the NOVANA dataset and we only look at year 2014. Click on this [link](data_6230.html) for further details on how this subset was created.

```{r}
cover <- read.csv("data/cover_data_6230_year2014.csv")

freq <- read.csv("data/frekvens_data_6230_year2014.csv")

abiotiske <- read.csv("data/abiotiske_data_6230_year2014.csv")

```
<br>
We apply the method function to the cover and freq datasets. 
```{r message=FALSE, warning=FALSE}
result <- bayesian_method(cover, freq, n = 16, 4)
```

```{r, include=FALSE}
beta_fit <- beta(cover, freq)
a <- as.numeric(beta_fit[beta_fit$species == "X116",]$a)
b <- as.numeric(beta_fit[beta_fit$species == "X116",]$b)
```

To get a sense of what the method has done, first we have plotted the prior distribution for the specie = "X116"

```{r}
curve(dbeta(x,a,b), xlab = "relative cover")
```

Below is histograms of the specie = "X116" in the observed cover dataset and in the updated abundance dataset where the above prior distrbution is used.
```{r}
hist(cover[,"X116"], probability = TRUE, xlab = "Number of pins", main = "Histogram of specie X116 in the observed cover dataset")
```

```{r}
hist(result[,"X116"], probability = TRUE, xlab = "Number of pins", main = "Histogram of specie X116 in the updated dataset")
```

<h1>Plot </h1>
As described on [link](Diversity.html) we can calculate different diveristy measures using the updated abundance data. In the below we have calculated the shannon index each plot in both the observed cover dataset and in the updated dataset:

```{r}
observed_shannon <- shannon(cover)
```


```{r}
updated_shannon <- shannon(result)
```


We plot the shannon indexes against each other
```{r, echo = FALSE}
ggplot(data = result, mapping = aes(x = observed_shannon, y = updated_shannon))+
  geom_point() +
  geom_abline(color = "blue") +
  ylab("Updated shannon index") +
  xlab("Observed shannon index") +
  ggtitle("Our shannon index compared with the observed shannon index")
```

It can be seen in this example, that the shannon index for the updated dataset gives a slightly higher value for most plots. Noticeably are the biggest difference among plots with a shannon value of 0 using the simpler method, that now have a value closer to 1 with our way of calculating the shannon index.


<br>
<br>

An interesting aspect of the data we work with in this example is how the diversity of a plot is correlated with the pH value of the soil at the plot. The function  [plot_index_gradient](gradient_plot.html) is used to visualize how different ways of calculating diversities compare against each other when mapped against the pH values. 
<br>
First we see the shannon index from the updated dataset relates to the pH values


```{r}
result$pH <- as.numeric(abiotiske$pHjord)
result$observed_shannon <- observed_shannon
result$updated_shannon <- updated_shannon
result$hill_shannon <- exp(result$updated_shannon)
result$species_richness <- species_richness(freq, species_indicator = "X")
```


```{r}
ggplot(result, aes(pH, updated_shannon))+
  geom_point()+
  xlim(3,6.5)+
  theme_classic()+
  labs(x = "pH value of the soil", y = "Updated shannon index")+
  ggtitle("Shannon index and pH value of soil at each plot")

```

Next we show how different diversity measures relates to the pH value

```{r, message=FALSE, warning=FALSE}
plot_index_gradient(result ,c("updated_shannon", "observed_shannon","species_richness", "hill_shannon"), "pH", gradient_text = "The PH value of the soil for each plot")
```













