---
title: "A bayesian method for better plant diverity estimates"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---
```{r, include=FALSE}
library(tidyverse)
library(kableExtra)
```

<h1>Introduction</h1>
Many methods for estimating the abundance of species within an area has been proposed in the litterature. In this vignette we will propose a new method with this goal where we combine cover data and prensence/absence data. We will be working with a subset of the NOVANA dataset as an example to show how this new method can be applied.
<br><br>
The NOVANA dataset consists of several parts. We will be using data of cover and presence/absence of species within each plot and also further data about each plot. 
<br>
Cover data of species in a given area can be done in many ways. In the NOVANA dataset the pin-point method is used to estimate the abundance of species in a plot. In the center of each plot $n$ pins is placed and the number of pins each specie is hit by is recorded. In the NOVANA dataset $n$ = 16 pins are used. As an example the below dataframe shows what cover data could look like
```{r, include=FALSE}
Cover <- data.frame(   a1 = c(2,0,5,0,1,0,16),
                       a2 = c(0,0,0,2,0,1,10),
                       a3 = c(0,0,1,1,16,0,2),
                       a4 = c(1,0,10,9,0,0,7),
                       a5 = c(0,0,0,3,0,0,0),
                       a6 = c(3,0,11,0,14,1,13),
                       a7 = c(0,0,7,4,0,1,0),
                       a8 = c(0,0,16,9,0,0,9),
                       a9 = c(2,0,0,1,11,0,5),
                       a10 = c(0,0,0,1,13,0,0))


rownames(Cover) <- c("Plot 1", "Plot 2", "Plot 3", "Plot 4", "Plot 5", "Plot 6", "Plot 7")
colnames(Cover) <- c("Specie 1", "Specie 2", "Specie 3", "Specie 4", "Specie 5", "Specie 6", "Specie 7", "Specie 8", "Specie 9", "Specie 10")
```

```{r}
kable(Cover, "html") %>% kable_styling()
```

The other part we will use to estimate the abundance of species in each plot is the precense/absence data. Within each plot it is recorded whether a specie is to be found somewhere in the plot. This is indicated by a binary variable with found = 1. Below is an example of what this data could look like
```{r, include=FALSE}
PersenceAbsence <- data.frame(   a1 = c(1,0,1,0,1,0,1),
                                 a2 = c(1,1,0,1,0,1,1),
                                 a3 = c(0,0,1,1,1,0,1),
                                 a4 = c(1,0,1,1,0,0,1),
                                 a5 = c(0,0,0,1,0,0,1),
                                 a6 = c(1,0,1,0,1,1,1),
                                 a7 = c(1,0,1,1,0,1,0),
                                 a8 = c(0,0,1,1,0,0,1),
                                 a9 = c(1,0,0,1,1,0,1),
                                 a10 = c(0,0,0,1,1,0,1))


rownames(PersenceAbsence) <- c("Plot 1", "Plot 2", "Plot 3", "Plot 4", "Plot 5", "Plot 6", "Plot 7")
colnames(PersenceAbsence) <- c("Specie 1", "Specie 2", "Specie 3", "Specie 4", "Specie 5", "Specie 6", "Specie 7", "Specie 8", "Specie 9", "Specie 10")
```


```{r}
kable(PersenceAbsence, "html") %>% kable_styling()
```

Here we also note that a specie can have a 1 in the presence/absence data but a 0 in the cover data as it is also the case in the above examples. We will make use of this property of the data in our method.

<h1>Theoretical discussion </h1>

The plant diversity is often calculated by only using the presence/absence data. However, in this method we loose a lot of valuable information from the cover data. In the following we propose a method that obtain an estimate of a specie's abundance in each plot that takes into account the spatial aggregation of the species which we can estimate using the cover data. This should also lead to better diversity estimates. We will do this by applying a bayesian framework and use a beta-binomial model. From the general setup in bayesian statistics we get that
$$
P(\theta|Y=y) \propto P(Y=y|\theta)\cdot \pi(\theta)
$$
where $P(Y=y|\theta)$ is the likelihood function and $\pi(\theta)$ is the prior distribution. We are interested in the posterior distribution $P(\theta|Y=y)$ which we can use to estimate the abundance of a specie in a plot. 
<br><br>
In the cover data each specie is hit by $y$ pins out of $n$ possible in each plot. A natural distribution to model this is the binomial distribution
$$
P(Y = y| \theta) = {n \choose y}\theta^y(1-\theta)^{n-y}
$$
<br>
We use the binomial model as the likelihood function in the bayesian framework that we apply.
<br>
We can create a prior distribution $\pi(\theta)$ for each specie by using both cover and presence/absence data. This is really the key idea in this method because this makes it possible to get a prior belief of a species spatial aggregation in a plot when we know that the specie is present.
<br><br>
The beta model is a conjugacy prior for the binomial model and, thus, is ideal to use. The beta-distribution is given by 
$$
\pi(\theta) = \text{beta}(\alpha, \beta) = \frac{\Gamma (\alpha+\beta)}{\Gamma (\alpha) \Gamma(\beta)}\theta^{\alpha -1}(1-\theta)^{\beta-1}
$$
When working with the bayesian framework, one has to decide on the paramters $\alpha$ and $\beta$. However, this can be difficult and instead we apply the empirical bayes method: For each specie we fit a beta-prior using all the cover data where the corresponding presence/absence data is 1. 
<br><br>
From the likelihood function and the prior distribution we get that the posterior will be an updated beta distribution for each specie in each plot
$$
P(\theta | Y = y) \sim Beta(\alpha + y, \beta + n-y)
$$
When the posterior distribution is found for a specie in a given plot, we use the mean posterior as a measure of the abundance for the specie in the plot. <br>
The mean posterior for a beta-distribution is given by:
$$
E\Big(P(\theta|Y =y)\Big) = \frac{\alpha + y}{\alpha + y+\beta + n-y}
$$
From the above theoretical discussion we get a new dataset with updated abundances for each specie in each plot. The next step for users of this method will be, as an xample, to apply some diversity measures for the plots in the new updated dataset. Some of commonly used diversity measures are described in [link](Shannon.html) and will be used in the examples below.
<h1> Implementation </h1>
```{r, include=FALSE}
source("code/function_shannon.R")
source("code/function.R")
```
The above theoretical discussions of the beta-binomial model are implemented in the function "bayesian_method". 

**Usage:**

bayesian_method(cover_data, presence_absence_data, n = 16 , remove_column = 0)

**Arguments:**

* **Cover_data**: Data frame with plots as rows and species as columns. Each cell in the data frame contains cover data for a given specie in a given plot.
* **presence_absence_data**: Data frame with plots as rows and species as columns. Each cell in the data frame contains a 1 if the specie is present in the plot or a 0 if the specie is absent in the plot.
* **n**: An integer. A number each cell in the cover data gets divided by if the cover data is not in the interval $[0,1]$. In the NOVANA dataset n = 16 secures this property. If the cover data is already on the interval $[0,1]$, one can just let n = 1.  
* **remove_column**: An integer. In case the first columns in the cover_data and presence_absence_data contain data not about the species but rather information on the plots, these columns have to be removed. The numbers of columns to be removed must be the same in the cover_data and presence/absence_data. The removed columns will be included again in the  output. In the NOVANA dataset the first columns are "site", "plot", year" so remove_column = 3 when this datasat is applied.

**Output** 

* The function returns a dataframe, similar to the cover data, but now containing new abundance data calculated using the method descripted above. 

For further specifications of the function click [link](bayesian_method_describtion.html)

<br>

<h1>First example </h1>
In the this first example we only work with data including the terhabtype = 6230 from the NOVANA dataset and we only look at year 2014. Click on this [link](data_6230.html) for further details on how this subset was created.
```{r}
cover <- read.csv("data/cover_data_6230_year2014.csv")
freq <- read.csv("data/frekvens_data_6230_year2014.csv")
abiotiske <- read.csv("data/abiotiske_data_6230_year2014.csv")
```
<br>
We apply the method function to the cover and freq datasets and obtain a "result" dataset with new abundace values for each species in each plot. 
```{r message=FALSE, warning=FALSE}
result <- bayesian_method(cover, freq, n = 16, 3)
```

<h3>Species level </h3>
```{r, include=FALSE}
beta_fit <- beta(cover, freq)
a <- as.numeric(beta_fit[beta_fit$species == "X116",]$a)
b <- as.numeric(beta_fit[beta_fit$species == "X116",]$b)
```
To get a sense of what the method has done, we plot the prior distribution of the species "Agrostis capillaris"
```{r}
curve(dbeta(x,a,b), xlab = "relative cover", main = "Prior distribution of Agrostis capillaris")
```
It is seen that there it high probability of being either around 0 or around 1. This is not only for this species but a common prior distribution for the species. This is because a lot of species are grouped in a way, such that if they are present, they will be present a lot, or else they are not present at all. 
<br><br>
Next we want to give a sense of how the the distributions of the number of pins the species Agrostis capillaris is hit by throughout all plots in the datasets change. First, we see the distribution of pins that the species is hit by in the observed cover data
```{r}
hist(cover[,"X116"], probability = TRUE, xlab = "Number of pins", main = "Histogram of the species Agrostis capillaris in the observed cover dataset")
```
It is seen that the species is not observed in a lot of plots, therefore the bar for 0 is quite high compared to the other bars. It is also seen that the other bars are somewhat similar. However, the bars get a bit higher towards the higher number of pins. This is in line with the grouping of a species described above. 
<br><br>
Below we plot the distributions of pins the species is hit by in the new "result" dataset
```{r}
hist(result[,"X116"], probability = TRUE, xlab = "Number of pins", main = "Histogram of the species Agrostis capillaris in the new dataset")
```
Here it is seen that bars from 1-16 are in general a bit higher than in the observed cover data. ... 
<br><br>
<h3>Plot level</h3>
We can also see some of the effects the method has had on the data at plot level. From the [Diversity](Diversity.html) page we have defined the Hill diversity as 
$$
\text{Hill diversity = }\left( \sum_{i=1}^{S} p_i (r_i)^{l}\right)^{1/l}
$$
In the below, we look at some plot in the dataset. We see how the diversity of the plot changes as a function of the exponent $l$. We do this for the plot in both the observed cover dataset and the new "result" dataset. This can help to give a feeling for which aspects of the the plot our method has changed. For $l = -1$ we have the Hill Simpson diversity, for $l = 0$ we have the Hill Shannon diversity and for $l = 1$ we have the species richness.
<br><br>



For plot 1
```{r, warning=FALSE}
different_diversities(cover, result, plot = 1,  remove_column = 3)
```

For plot 100
```{r, warning=FALSE}
different_diversities(cover, result, plot = 100, remove_column = 3)
```

In both plots we see how the diversity estimates drift apart when $l$ increases. For values of $l$ below zero we emphasize common species higher while values of $l$ above zero emphasize rare species higer.
<br>

To get an idea for how the whole data set is we also plot the average for multiple plots.


```{r, warning=FALSE}
different_diversities2(cover[1:5,], result[1:5,],  remove_column = 3)
```


...

<h3> Different diversity estimates compared </h3>

We can also compare how different diversity estimates differ throughout all the plots in the observed cover data and the new dataset respectively. All the used diversity estimates are defined at [link](Diversity.html).
In the below we have calculated the shannon index each plot in both the observed cover dataset and in the "result" dataset

```{r}
observed_shannon <- shannon(cover, 3)
```

```{r}
updated_shannon <- shannon(result, 3)
```

We plot the shannon indexes against each other
```{r, echo = FALSE}
ggplot(mapping = aes(x = observed_shannon, y = updated_shannon))+
  geom_point() +
  geom_abline(color = "blue") +
  ylab("Updated shannon index") +
  xlab("Observed shannon index") +
  ggtitle("Our shannon index compared with the observed shannon index")
```
It can be seen in this example, that the shannon index for the updated dataset gives a slightly higher value for most plots. Noticeably are the biggest difference among plots with a shannon value of 0 using the simpler method, that now have a value closer to 1 with our way of calculating the shannon index.
<br>
<br>
Another often used diversity estimate is the simpson's index. We create a similar plot as above
```{r}
observed_simpson <- simpson(cover, 3)
```

```{r}
updated_simpson <- simpson(result, 3)
```

```{r, echo = FALSE}
ggplot(mapping = aes(x = observed_simpson, y = updated_simpson))+
  geom_point() +
  geom_abline(color = "blue") +
  ylab("Updated simpson index") +
  xlab("Observed simpson index") +
  ggtitle("Our simpson index compared with the observed shannon index")
```

Small values of simpson index indicate a high diveristy for a plot. It is worth to notice that for plots with high diversity our method does not seem to change the data as much as it did in the plot of the shannon indexes. 
<br><br>
...

<h3>Species diversity and pH value </h3>

An interesting aspect of the data we work with in this example is how the diversity of a plot is correlated with the pH value of the soil at the plot. 
<br>
First we see the shannon index from the updated dataset relates to the pH values

```{r, warning = FALSE}
ggplot(mapping = aes(x = as.numeric(abiotiske$pHjord), y = updated_shannon)) +
  geom_point()+
  xlim(3,6.5)+
  theme_classic()+
  labs(x = "pH value of the soil", y = "Updated shannon index")+
  ggtitle("Shannon index and pH value of soil at each plot")
```

Next we show how different diversity measures relates to the pH value. The function  [plot_index_gradient](gradient_plot.html) is used to visualize how different ways of calculating diversities compare against each other when mapped against the pH values. 

```{r, warning = FALSE, include = FALSE}
data_for_plot <- as.data.frame(as.numeric(abiotiske$pHjord))
colnames(data_for_plot) <- "pH"
data_for_plot$'Shannon index' <- shannon(result, 3)
data_for_plot$'Hill Shannon' <- hill_shannon(result, 3)
data_for_plot$'Species richness' <- species_richness(freq, remove_column = 3)
data_for_plot$'Simpson index' <- simpson(result, 3)
data_for_plot$'Hill Simpson' <- hill_simpson(result, 3)
```

```{r, message=FALSE, warning=FALSE}
plot_diversity(data_for_plot ,c("Hill Shannon","Species richness", "Hill Simpson"), "pH", description = "The PH value of the soil for each plot")
```

Comments ... 

<h1> Second example  </h1>
In the second example we look at terhabtype = 6210 in year 2009 from the NOVANA dataset.
```{r}
cover_6210 <- read.csv("data/cover_data_6210_year2009.csv")
freq_6210 <- read.csv("data/frekvens_data_6210_year2009.csv")
abiotiske_6210 <- read.csv("data/abiotiske_data_6210_year2009.csv")
```



With this data, we are especially interested in how the diversity change when plotted against different values of nitrogen in the soil. It is known that increased nitrogen supply can cause changes to the species composition. 
<br><br>
We apply our function to the data. 
```{r}
new_data <- bayesian_method(cover_6210, freq_6210, n = 16, remove_column = 3)
```


```{r, warning = FALSE, include = FALSE}
data_for_plot <- as.data.frame(as.numeric(abiotiske_6210$Nijord))
colnames(data_for_plot) <- "N"
data_for_plot$'Shannon index' <- shannon(new_data, 3)
data_for_plot$'Hill Shannon' <- hill_shannon(new_data, 3)
data_for_plot$'Species richness' <- species_richness(freq_6210, remove_column = 3)
data_for_plot$'Simpson index' <- simpson(new_data, 3)
data_for_plot$'Hill Simpson' <- hill_simpson(new_data, 3)
```

We map the shannon index of the new dataset against the nitrogen value of the soil 
```{r, echo = FALSE, warning=FALSE}
ggplot(data_for_plot, mapping = aes(x = N, y = `Shannon index`))+
  geom_point() +
  ylab("Shannon index") +
  xlab("Nitrogen in soil") +
  ggtitle("Our shannon index compared to nitrogen value in the soil")
```

```{r, warning=FALSE}
ggplot(data_for_plot, mapping = aes(x = N, y = `Hill Simpson`))+
  geom_point() +
  ylab("Hill Simpson") +
  xlab("Nitrogen in soil") +
  ggtitle("Hill Simpson compared to nitrogen value in the soil")

```


```{r, warning=FALSE}
ggplot(data_for_plot, mapping = aes(x = N, y = `Species richness`))+
  geom_point() +
  ylab("Species Richness") +
  xlab("Nitrogen in soil") +
  ggtitle("Species richness compared to nitrogen value in the soil")

```

