---
title: "Main page"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---
```{r, include=FALSE}
library(tidyverse)
library(kableExtra)
```

<h1>Introduction</h1>
Many methods for estimating the abundance of species within an area has been proposed in the litterature. In this vignette we will propose a new method with this goal where we combine cover data and prensence/absence data. We will be working with a subset of the NOVANA dataset as an example to show how this new method can be applied.
<br><br>
The NOVANA dataset consists of several parts. We will be using data of cover and presence/absence of species within each plot and also further data about each plot. 
<br>
Cover data of species in a given area can be done in many ways. In the NOVANA dataset the pin-point method is used to estimate the abundance of species in a plot. In the center of each plot $n$ pins is placed and the number of pins each specie is hit by is recorded. In the NOVANA dataset 16 pins are used. As an example the below dataframe shows what cover data could look like
```{r, include=FALSE}
Cover <- data.frame(   a1 = c(2,0,5,0,1,0,16),
                       a2 = c(0,0,0,2,0,1,10),
                       a3 = c(0,0,1,1,16,0,2),
                       a4 = c(1,0,10,9,0,0,7),
                       a5 = c(0,0,0,3,0,0,0),
                       a6 = c(3,0,11,0,14,1,13),
                       a7 = c(0,0,7,4,0,1,0),
                       a8 = c(0,0,16,9,0,0,9),
                       a9 = c(2,0,0,1,11,0,5),
                       a10 = c(0,0,0,1,13,0,0))


rownames(Cover) <- c("Plot 1", "Plot 2", "Plot 3", "Plot 4", "Plot 5", "Plot 6", "Plot 7")
colnames(Cover) <- c("Specie 1", "Specie 2", "Specie 3", "Specie 4", "Specie 5", "Specie 6", "Specie 7", "Specie 8", "Specie 9", "Specie 10")
```

```{r}
kable(Cover, "html") %>% kable_styling()
```

The other part we will use to estimate the abundance of species in each plot is the precense/absence data. Within each plot it is recorded whether a specie is to be found somewhere in the plot. This is indicated by a binary variable with found = 1. Below is an example of what this data could look like
```{r, include=FALSE}
PersenceAbsence <- data.frame(   a1 = c(1,0,1,0,1,0,1),
                                 a2 = c(1,1,0,1,0,1,1),
                                 a3 = c(0,0,1,1,1,0,1),
                                 a4 = c(1,0,1,1,0,0,1),
                                 a5 = c(0,0,0,1,0,0,1),
                                 a6 = c(1,0,1,0,1,1,1),
                                 a7 = c(1,0,1,1,0,1,0),
                                 a8 = c(0,0,1,1,0,0,1),
                                 a9 = c(1,0,0,1,1,0,1),
                                 a10 = c(0,0,0,1,1,0,1))


rownames(PersenceAbsence) <- c("Plot 1", "Plot 2", "Plot 3", "Plot 4", "Plot 5", "Plot 6", "Plot 7")
colnames(PersenceAbsence) <- c("Specie 1", "Specie 2", "Specie 3", "Specie 4", "Specie 5", "Specie 6", "Specie 7", "Specie 8", "Specie 9", "Specie 10")
```


```{r}
kable(PersenceAbsence, "html") %>% kable_styling()
```

Here we also note that a specie can have a 1 in the presence/absence data but a 0 in the cover data as it is also the case in the above examples. We will make use of this property of the data in our method.
<br><br>
The plant diversity is often calculated by only using the cover data. However, in this method we loose a lot of valuable information from the presence/absence data. In the following we propose a method that obtain an estimate of a specie's abundance in each plot that takes into account the spatial aggregation of the species. This should also lead to better diversity estimates. We will do this by applying a bayesian framework and use a beta-binomial model.
<br><br>
In the cover data each specie is hit by $y$ pins out of $n$ possible. To model this we apply the binomial distribution
$$
P(Y = y| \theta) = {n \choose y}\theta^y(1-\theta)^{n-y}
$$
<br>
In the bayesion framework the binomial model is used as a likelihood function when modeling the posterior distribution for a specie's abundance in each plot.
<br>
We can create a prior distribution $\theta$ for each specie by using both cover and presence/absence data. The beta model is a conjugacy prior for the binomial model and, thus, is ideal to use. The beta-model is given by 
$$
f(\theta) = \frac{\Gamma (\alpha+\beta)}{\Gamma (\alpha) \Gamma(\beta)}\theta^{\alpha -1}(1-\theta)^{\beta-1}
$$
For each specie we fit a beta-prior using all the cover data where the corresponding presence/absence data is 1.  
<br><br>
Using the above we can form a posterior, that will be an updated beta distribution for each specie in each plot
$$
\theta | (Y = y) \sim Beta(\alpha + y, \beta + n-y)
$$
When the posterior distribution is found for a specie in a given plot, we use the mean posterior as a measure of the abundance for the specie in the plot. <br>

The mean posterior for a beta-distribution is given by:

$$
\frac{\alpha + y}{\alpha + y+\beta + n-y}
$$


<h1>An example </h1>

We work with data only including the terhabtype = 6230 from the NOVANA dataset and we only look at year 2014. Click on this [link](data_6230.html) for further details on how this subset was created.

```{r}
cover <- read.csv("data/cover_data_6230_year2014.csv")

freq <- read.csv("data/frekvens_data_6230_year2014.csv")

abiotiske <- read.csv("data/abiotiske_data_6230_year2014.csv")

```
<br>

```{r, include=FALSE}
source("code/function_shannon.R")
```

The above theoretical discussions of the beta-binomial model are implemented in the function shanon_index_v2. For further specifications of the function click [link](function_shannon.html)



Apply our shannon function on the data
```{r, message=FALSE, warning=FALSE}
result <- shanon_index_v2(cover, freq)
```



```{r}
kable(result, "html") %>%
    kable_styling() %>%
    scroll_box(width = "800px", height = "400px")
```
<br><br>

<h1>Plot </h1>

Look at connection between pH value of the soil and the shannon index for the plot.

```{r}
result$pH <- abiotiske$pHjord
plot_result <- result[!(result$pH == "mv"),]
plot_result$pH <- as.numeric(plot_result$pH)

ggplot(plot_result, aes(pH, shannon))+
  geom_point()+
  xlim(3,6.5)+
  theme_classic()+
  labs(x = "pH value of the soil", y = "Shannon index")+
  ggtitle("Shannon index and pH value of soil at each plot")

```














